---
title: "Homework 3"
author: Vera Xu
output: github_document
---

This is my solution to HW3.

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r, message = FALSE, warning = FALSE}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r, message = FALSE, warning = FALSE}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r, message = FALSE, warning = FALSE}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r, message = FALSE, warning = FALSE}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

### Question 2

* Load and tidy the data using the following code chunk:

```{r, message = FALSE, warning = FALSE}
p2_df = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
   activity_1:activity_1440,
   names_prefix = "activity_",
   names_to = "min",
   values_to = "activity"
   ) %>%
   mutate(
    weekday = ifelse(day %in% c("Saturday", "Sunday"), FALSE, TRUE),
    min = as.numeric(min),
    dow = recode_factor(day,
                        "Sunday" = 0,
                        "Monday" = 1, 
                        "Tuesday" = 2, 
                        "Wednesday" = 3,
                        "Thursday" = 4, 
                        "Friday" = 5, 
                        "Saturday" = 6)
    )%>%
  select(-day_id)%>%
  order(c(week, dow))
```

The resulting dataset contains `r ncol(p2_df)` variables, including week, day of the week, weekday (yes or no), minute of the day, activity. There are `r nrow(p2_df)` observations in total.


* Compute total activity over the day

```{r, message = FALSE, warning = FALSE}
p2_df %>%
  group_by(week, day) %>% 
  summarize(total_activity = sum(activity))%>% 
  knitr::kable(digits = 2)
```

** It seems that this person tend to have a stable level of activity over the weekdays, with some extremely low or high levels on certain weekends.
** It also seems like this person is more active during week 3.


* Activity level plot

```{r, message = FALSE, warning = FALSE}
p2_df %>%
  ggplot(aes(x = min, y = activity)) + 
  geom_line(aes(colour = day), alpha = 0.5)
```

** It seems that this person is active during daytime and not active over the night.



### Question 3

```{r}
  data("ny_noaa")
```

```{r}
  ny_noaa %>% 
  separate(date, c("month", "day", "year"), sep = "-") %>%
  summarize(prcp)

```

* Make a plot showing the average max temperature in January and in July

```{r}
ny_noaa %>% 
  separate(date, c("year", "month", "day"), sep = "-") %>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    day = as.numeric(day),
    tmin = as.numeric(tmin)*0.1,
    tmax = as.numeric(tmax)*0.1) %>%
  filter(month == 1 | month == 7) %>%
  

```

*