---
title: "Homework 2"
author: Vera Xu
output: github_document
---

This is my solution to HW2.

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1 as demonstrated by Jeff during lecture

First, define a path to the dataset. 

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2

#### Importing, Tidying and description of the NYC Transit data

Import the "NYC_Transit_Subway_Entrance_And_Exit_Data.csv" dataset in the code chunk below:

```{r import NYC Transit data, message = FALSE, warning = FALSE}
p2_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select (line, station_name, station_latitude, station_longitude, 
          route1:route11,  entrance_type, entry, vending, ada) %>%
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE),
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
    ) %>%
  pivot_longer(
   route1:route11,
    values_to = "route_name"
   ) %>%
  select(-name)
```


This dataset contains NYC Transit information. After cleaning and tidying, it contains the following variables: line, station name, station latitude, station longitude, routes served, entry, vending, entrance type, and ADA compliance status. 
The steps I've used to clean the data so far are:

* Clean column names
* Select the columns as required
* Convert the entry variable from character to a logical variable and convert all route variables as character
* Keep distinct rows only
* Transform the data into long from, values of route1 to route11 variables are stored as route_name, NA values are dropped

After the cleaning and tidying steps above, the cleaned dataset has a dimension of (`r nrow(p2_df)` rows * `r ncol(p2_df)` columns).

The raw dataset is not tidy as there are repeated information, and varialbles used to annotate routes is in the wide form.

#### Answers to Problem 2 question list

* There are `r distinct(p2_df, station_name, line) %>% nrow()` distinct stations in this dataset.
* There are `r filter(p2_df, ada == 'TRUE') %>% distinct(station_name, line) %>% nrow()` ADA compliant stations.
* There are `r (filter(p2_df, vending == "NO" & entry == TRUE) %>% nrow() / filter(p2_df, vending == "NO") %>% nrow()) %>% round(digits = 4) * 100`% station entrances/exits without vending allow entrance.
 
#### Route A specific stations

* The route number and route name are already distinct when tidying the imported data in the process above.
* There are `r filter(p2_df, route_name == 'A') %>% distinct(station_name, line) %>% nrow()` distinct stations serve the A train. Of these stations, `r filter(p2_df, route_name == 'A' & ada == 'TRUE') %>% distinct(station_name, line) %>% nrow()` are ADA compliant.


## Problem 3

Read in and clean the pols-month dataset in the following code chunk:

```{r, message = FALSE, warning = FALSE}
p3_df1 = 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, c("year", "month", "day"), sep = "-", convert = TRUE) %>%
  mutate(
    month = recode(month,
                   "1" = "January",
                   "2" = "Feburary",
                   "3" = "March",
                   "4" = "April",
                   "5" = "May",
                   "6" = "June",
                   "7" = "July",
                   "8" = "August",
                   "9" = "September",
                   "10" = "October",
                   "11" = "November",
                   "12" = "December"),
    president = ifelse(prez_gop == 0, "Demographic", "Republican")
    ) %>%
  select(-prez_gop, -prez_dem,-day)
```

Read in and clean the snp dataset in the following code chunk:

```{r, message = FALSE, warning = FALSE}
p3_df2 = 
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, c("month", "day", "year"), sep = "/", convert = TRUE) %>%
  mutate(month = recode(month,
                   "1" = "January",
                   "2" = "Feburary",
                   "3" = "March",
                   "4" = "April",
                   "5" = "May",
                   "6" = "June",
                   "7" = "July",
                   "8" = "August",
                   "9" = "September",
                   "10" = "October",
                   "11" = "November",
                   "12" = "December")) %>%
  select(year, month, close)
```


Read in and clean the unemployment dataset in the following code chunk:

```{r, message = FALSE, warning = FALSE}
p3_df3 = 
  read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(c("Jan","Feb", "Mar", "Apr", "May","Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
               names_to = "month",
               values_to = "rate") %>%
  mutate(month = recode(month,
                   "Jan" = "January",
                   "Feb" = "Feburary",
                   "Mar" = "March",
                   "Apr" = "April",
                   "Jun" = "June",
                   "Jul" = "July", 
                   "Aug" = "August", 
                   "Sep" = "September",
                   "Oct" = "October", 
                   "Nov" = "November", 
                   "Dec" = "December")) %>%
  rename_with(tolower)
```

Join the datasets by merging snp into pols, and merging unemployment into the result:

```{r, message = FALSE, warning = FALSE}
p3_join_result = inner_join(p3_df1, p3_df2, by = c("year", "month")) %>%
  inner_join(p3_df3, by = c("year", "month"))
```

#### Description of the read-in datasets

The “pols-month” dataset, after tidied, contains information related to the number of national politicians who are democratic or republican at given year and months:

* There are `r ncol(p3_df1)` columns in total. Key variables are year, month and president's political position (variable name "president", coded as republican or democratic).
* There are `r nrow(p3_df1)` rows in total.
* The covered year range in this dataset is from `r min(p3_df1[,1])` to `r max(p3_df1[,1])`.

The “snp” dataset, after tidied, contains information related to Standard & Poor’s stock market index at given year and months:

* There are `r ncol(p3_df2)` columns in total. Variables are year, month and closing values of the S&P stock index.
* There are `r nrow(p3_df2)` rows in total.
* The covered year range in this dataset is from `r min(p3_df2[,1])` to `r max(p3_df2[,1])`.

The “unemployment” dataset, after tidied, contains information related to Standard & Poor’s stock market index at given year and months:

* There are `r ncol(p3_df3)` columns in total. Variables are year, month and percentage of unemployment.
* There are `r nrow(p3_df3)` rows in total.
* The covered year range in this dataset is from `r min(p3_df3[,1])` to `r max(p3_df3[,1])`.

#### Description of the merged dataset

* The three dayasets are merged by year and month using a inner join method. The joined dataset has a dimention of (`r nrow(p3_join_result)` rows * `r ncol(p3_join_result)` columns). 
* The covered year range in the merged data is from `r min(p3_join_result[,1])` to `r max(p3_join_result[,1])`. 
* The columns in the joined dataset are: `r ls(p3_join_result, sorted = FALSE)`.
* Key variables including year, month, president (democratic or republican), closing values of the S&P stock index, percentage of unemployment corresponding to that specific year and month.
